\mythesistitle

### 2011-11-20 src/synopsis.tex
- corrected the synopsis template usage. - deleted chapters. made sections - basic org done. - the main changes and final! changes are pending.

\documentclass[MS,]{iitmdiss} 
\usepackage[usenames,dvipsnames]{color}

@@

\nocite{*}
\maketitle

\begin{singlespace}
\tableofcontents
\thispagestyle{empty}
\listoftables
\addcontentsline{toc}{chapter}{LIST OF TABLES}
\listoffigures
\addcontentsline{toc}{chapter}{LIST OF FIGURES}
\end{singlespace}

@@

\chapter{Introduction}
\label{chap:intro}

\tnote{give some basic gyan about the problem itself? or screw it move to
section directly}

\section{Illustration of the problem}
\label{sec:problem}

@@

Let $U$ and $\cF$ represent the set of students and the set of study
groups respectively and the integers $n$ and $m$ denote the total
number students and study groups respectively. In relation to this
example, these are defined as follows. Also given below is the study group
allocation to students. \tnote{use tables}
\tnote{use other font for elements of sets}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{ll}
    $U $&$= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
    $\cF $&$= \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
    $\xLLL $&$= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
    $\xGGG $&$= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
    $\xBBB $&$= \{\xSn,\;  \xPi,\;  \xWo \}$\\
    $\xTTT $&$= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
    $n $&$= |U| = 11$\\
    $m $&$= |\cF| = 4$      
  \end{tabular}
  \caption{Students and study groups in \WSI}
  \label{tab:wsigroups}
\end{table}

@@

\begin{figure}[htbp]
  \centering
%  \includegraphics[scale=0.5]{../img/2_infinite_loop.png}
  \includegraphics[scale=0.5]{../img/2_infinite_loop_BTWF.png}
  \caption{{\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
  \label{fig:streetmappath}
\end{figure}

@@

As a special case, suppose all the apartments are on the same
street or if they are all lined up on a single path, the street map
becomes a tree that is just a path. Then the problem becomes what is
called an {\em interval assignment problem}. The idea of interval
assignment may not be obvious here; hence to see this, consider a
different problem in {\WSI} where the classes for these study groups
courses need to be scheduled during a day (or a week or any time
period). Each study group has a bunch of courses associated with it
some of which may be shared by two or more study groups. It is
mandatory that a student who is a member of a study group takes all
the courses associated with that group. There are slots during the day
for classes to be held and the problem is to allocate class slots to
courses such that all the classes of a study group are consecutive. It
is debatable if this will not hamper the attention span and memory
retention rate of the students but that is, regrettably, out of the
scope of this thesis. The parallels between this class allocation
problem and the accommodation problem can be seen as follows. The set
$U$ here, are the courses offered (say Course 101 {\coneohone}, Course
102 {\coneohtwo} and so on). In this variation of the problem, the
collection $\cF$ is the set of study groups but the study groups are
filled by course IDs (in place of students in the earlier
example). For instance, Course 101 is mandatory for all study groups
$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
only the $\xLLL$ group) and so on. The sequence of class slots for the
day (or week or any time period) is analogous to the street map in the
accommodation problem. It is quite obvious now why this version of the
problem (where the ``target graph'' is a path and not any
tree\tnote{\footnote{Allowing any tree in this example could be seen
    as a scenario where there are parallel classes. A node falling in
    the path between two other nodes would mean that the corresponding
    is scheduled between the other two.}}) is called an interval
assignment problem.

\tnote{TBD: BEGIN REVIEW FROM BELOW}

@@

\tnote[a]{THESIS: UNCOMMENT} \tnote{include the rest of the sections as
  the organization of the doc}
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
% Section~\ref{sec:feasible} documents the
% characterization of a feasible path labeling and finally,
% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
% find the tree path labeling of a given set system from a given
% $k$-subdivided tree.

Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
Chapter~\ref{chap:results} presents a summary of our results on the
tree path labeling problem.

@@

As descirbed earlier, a matrix with COP is one whose rows (columns)
can be rearranged so that the 1s in every column (row) are in
consecutive rows (columns).

@@

\cite{bl76} describes a linear time construction of a data structure
called PQ trees. This paper draws on the close relationship of COP
matrices to interval graphs. A PQ tree of a matrix is one that stores
all row (column) permutations of the matrix that give the COP orders
(there could be multiple) of the matrix. This is constructed using an
elaborate linear time procedure. Since this
result, several improvements have been made to the procedural
complexity of the algorithm itself and also derivative data structures
that are easier to construct. PQR trees is one such data structure
\cite{mm96}, \cite{mpt98}\tnote{also, mpt98 $==$ elaborate version of
  mm96}.  \cite{tm05} describes an improved version of PQR
trees. \tnote[a]{improvement in terms of what?}\cite{wlh02} describes
the simpler algorithm for COT. \tnote[a]{This result first apprearted
  in proc ISAAC 92} Hsu also invented PC trees \cite{wlh01} which is
claimed to be much easier to implement. \cite{nsnrs09} describes a
characterization of consecutive ones property solely based on the
cardinality properties of the set representations of the columns
(rows); every column (row) is equivalent to a set that has the row
(column) indices of the rows (columns) that have one entries in this
column (row). This is interesting and relevant, especially to this
thesis because it simplifies COT to a great degree. \tnote[a]{because
  it reduces the search space for the solution from set of all
  permutations of rows to set of all-- get this figured out!}

\cite{mcc04} describes a different approach to COT. While all previous
COT algorithms gave the COP order if the matrix has the property and
if not it exited stating negative, this algorithm gives an evidence by
way of a certificate of
matrix being devoid of COP thereby enabling a user to check the
algorithm's result. This is significant from an implementation
perspective because automated program verification is hard and manual
verification is more viable. Hence having a certificate
reinforces its credibility. Note that when the matrix has COP the COP
order is the certificate.  It may also be noted that the internal
machinery of this algorithm \annote[a] {is related to the weighted
  betweenness problem addressed } {in what way??} in \cite{co98}.
\tnote[a]{expand on the COP order graph creation and it having to be
  bipartite for M to have COP. and thus an odd cycle being an evidence
  of no COP.}

\tnote[a]{THESIS: where should this go?: \cite{jlm97} (application of
  PQ trees in graphics).}

The literature also has results on parallel COT algorithms --
\cite{as95} \cite{bs03} \tnote[a]{is this parallel?}
\cite{ly91}\tnote[a]{is this parallel?}

\tnote[a]{THESIS: add helly's theorem citation
  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}

{\bf Matrices with no COP}: So far we have been concerned about matrices that have the consecutive
ones property. However in real life applications, it is rare that data
sets represented by binary matrices have COP. Nevertheless, the COP is
not arbirary and there are several interesting problems when a matrix
does not have COP but is ``close'' to having COP or can be altered to
have COP.

-- sect 4.1 in \cite{d08phd} has many results surveyed. hardness results,
approx. results. results are usually for a class of matrices $(a,b)$
where number columns and rows are restriced to $a$ and $b$
\tnote[a]{check if it is the other way around}.\\
-- problem of flipping at most $k$ entries of $M$ to make it attain
COP. this is NP complete \cite{b75-phd}\\

The first step in this direction is to analyze matrices without
COP. \cite{at72} showed that a matrix that does not have COP have
certain forbidden substructures that prevent it from having
COP. Tucker classified these substructures into five classes of
submatrices. This result is presented in the context of AT-free
bipartite graphs \tnote[a]{THESIS: check this up. give details.} which
is equivalent to matrices with COP. A binary matrix is converted into
a bipartite graph as follows. The set of rows and the set of columns
form the two partitions of the graph. Each row node is adjacent to
those nodes that represent the columns that have ones in the
corresponding row. \cite{at72} proves that this bipartite graph has no
asteroidal triple if and only if the matrix has COP and goes on to
identify the forbidden substructures for these bipartite graphs. The
matrices corresponding to these substructures are the forbidden
submatrices. Once a matrix has been detected to not have COP (using
any of the COT algorithms mentioned earlier), it is naturally of
interest to find out the smallest\tnote[a]{define ``smallest''}
forbidden substructure. \cite{d08phd} discusses an exact algorithm
with time complexity \annote[a]{parameterized by the number of ones in
  the matrix}{check this.}.

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete. \tnote[a]{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hg02} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem. \tnote[a]{THESIS: elaborate - what are
  the results?}.

Another problem is to find the minumum number of entries in the matrix
that can be flipped to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to ones so that the
resulting matrix has COP. As mentioned earlier, this problem had been known
to be NP complete due to \cite{b75-phd}. \cite{v85} also proves, using
a reduction to the longest path problem, \tnote[a]{or is it a survey
  of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$
rows is NP complete. \tnote[a]{how is this different from booth's 75 result??} \tnote{[tag:appoximation,bitFlipping,matrixBreaking]}\\

\tnote[a]{THESIS: where should this go? \cite{tz04} (approx submatrix with COP sparse matrices)}

%\subsection{Further extensions of the COP problem}

\cite{jkckv04} \tnote{[tag:storage]} discusses the use of matrices
with ``near'' COP in the storage of very large databases.
The problem is that of $k$-run reordering of binary matrices which
they prove to be NP hard using a reduction from the hamiltonian path problem.
A connection of COP problem to the travelling salesman problem is also
introduced. \tnote[a]{THESIS? what does this mean? -- COP can be used as a tool to reorder $0.5T \le runs(M) \le T$}

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{k77}. \tnote[a]{THESIS: what is the reduction?}

\tnote[a]{THESIS: 
-- {\em other problems similar to COP}
  \begin{enumerate}
  \item \cite{ckl96} (ILP, circ ones, one drop)
  \item \cite{th98} (generalization of COP - minimax, biotonic column) Tucker
  \end{enumerate}
}

{\bf Motivation: } A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote[a]{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

Our results also have close connection to recognition of path graphs
and connection to path graph isomorphism.  Graphs which can be
represented as the intersection graph of paths in a tree are called
{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
be viewed as paths in a tree, if and only if the intersection graph of
$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
graphs since chordal graphs are combinatorially characterized as the
intersection graphs of subtrees of a tree.  Path graphs are well
studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
a graph is a path graph can be done in polynomial time by the results
of \cite{gav78,aas93}.  However, this is only a necessary condition
for our question.  \remove[a]{Indeed, our work provides a sufficient
  condition for a related problem: that is whether a path assignment
  to be feasible?}  On the other hand, path graph isomorphism is known
be isomorphism-complete; see for example \cite{kklv10}. Therefore,
\annote[a]{it is unlikely that we can solve the problem of finding
  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
\annote[a]{It is definitely interesting to classify the kinds of trees
  and hypergraphs for which feasible path labelings can be
}{ELABORATE.} found efficiently.  \annote[a]{These results would form
  a natural generalization of COP testing and interval graph
  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.

@@

\chapter{Summary of results}
\label{chap:results}

@@

%\item 
In the Section 4, \annote[a]{we initiate an exploration of}{WTF?}
finding feasible path labeling of set systems in a special kind of
tree which we call the $k$-subdivided star.  \annote[a]{This question
  is an attempt to generalize}{REWORD.} the problem of testing if a
matrix has the consecutive ones property.  However, we restrict the
hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
elements.  \annote[a]{In spite of this restricted case we consider, we
  believe that our results are of significant interest in
  understanding the nature of Graph Isomorphism}{REWORD.} which is
polynomial time solvable in interval graphs and is hard on path
graphs.
%\end{enumerate}

@@

\section{RESULTS}
% \section{Preliminaries} 
% \label{sec:prelims}

@@

\section{Characterization of Feasible Tree Path  Labelings} 
\label{sec:feasible} 
In this section we give an algorithmic characterization of a
feasibility of tree path labeling.
% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% and path system $\cP$ on the given tree $T$. We
% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
%   (ICPPL)} if it has the following properties.
Consider a path labeling $(\cF, \cl)$ on the
given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
  Preserving Path Labeling (ICPPL)} if it has the following
properties.

\begin{enumerate}[{\icpplpr\ }i. ]
\item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
  % \vspace{\topshrink}
\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
  for all distinct $S_1, S_2 \in \cF$
  % \vspace{\topshrink}
\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
  \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
\end{enumerate}

\noindent
The following lemma is useful in subsequent arguments. 
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\tnote[a]{THESIS: UNCOMMENT}
% \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
%   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
%   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
%   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
%   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
%   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
%   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
%   P_3)|$. Thus lemma is proved. \qed
% \end{proof}

>>>>

\documentclass[MS,synopsis]{iitmdiss} 
\usepackage[usenames,dvipsnames]{color}

@@

%\nocite{*}
\maketitle

% \tnote{TBD: DELETE IF WE RUN OUT OF SPACE}
% \begin{singlespace}
% \tableofcontents
% \thispagestyle{empty}
% \listoffigures
% \addcontentsline{toc}{chapter}{LIST OF FIGURES}
% \listoftables
% \addcontentsline{toc}{chapter}{LIST OF TABLES}
% \end{singlespace}

@@

\section{Introduction}
\label{sec:intro}

% \tnote{testing 1 2 3...}
% \tnote[bogus1]{testing 1 2 3...}
% \tnote[bogus2]{testing 1 2 3...}
% \tnote[bogus3]{testing 1 2 3...}
% \tnote[bogus4]{testing 1 2 3...}
% \tnote[bogus5]{testing 1 2 3...}
% \tnote[bogus6]{testing 1 2 3...}
% \tnote[bogus7]{testing 1 2 3...}
% \tnote[bogus8]{testing 1 2 3...}
% \tnote[TH]{testing 1 2 3...}

\subsection{Illustration of the problem}
\label{sec:problem}

@@

Let $U$ and $\cF$ represent the set of students and the set of study
groups respectively and the integers $n$ and $m$ denote the total
number students and study groups respectively. In relation to this
example, these are defined in Table~\ref{tab:wsigroups}. Also given
there is the study group allocation to students.

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{rcl}
    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
    $n $&$=$&$ |U| = 11$\\
    $m $&$=$&$ |\cF| = 4$      
  \end{tabular}
  \caption{Students and study groups in \WSI}
  \label{tab:wsigroups}
\end{table}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
  \caption{{\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
  \label{fig:streetmappath}
\end{figure}

@@

As a special case, suppose all the apartments are on the same street
or if they are all lined up on a single path, the street map becomes a
tree that is just a path. Then the problem becomes what is called an
{\em interval assignment problem}. The idea of interval assignment may
not be obvious here; hence to see this, consider a different problem
in {\WSI} where the classes for these study groups courses need to be
scheduled during a day (or a week or any time period). Each study
group has a bunch of courses associated with it some of which may be
shared by two or more study groups. It is mandatory that a student who
is a member of a study group takes all the courses associated with
that group. There are slots during the day for classes to be held and
the problem is to allocate class slots to courses such that all the
classes of a study group are consecutive. It is debatable if this will
not hamper the attention span and memory retention rate of the
students but that is, regrettably, out of the scope of this
thesis. The parallels between this class allocation problem and the
accommodation problem can be seen as follows. The set $U$ here, are
the courses offered (say Course 101 {\coneohone}, Course 102
{\coneohtwo} and so on). In this variation of the problem, the
collection $\cF$ is the set of study groups but the study groups are
filled by course IDs (in place of students in the earlier
example). For instance, Course 101 is mandatory for all study groups
$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
only the $\xLLL$ group) and so on. The sequence of class slots for the
day (or week or any time period) is analogous to the street map in the
accommodation problem. It is quite obvious now why this version of the
problem (where the ``target graph'' is a path and not any
tree\tnote[TH]{Allowing any tree in this example could be seen as a
  scenario where there are parallel classes. A node falling in the
  path between two other nodes would mean that the corresponding is
  scheduled between the other two.}) is called an interval assignment
problem.

@@

\tnote{add a note on the organization of the doc}
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
% Section~\ref{sec:feasible} documents the
% characterization of a feasible path labeling and finally,
% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
% find the tree path labeling of a given set system from a given
% $k$-subdivided tree.

Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
Section~\ref{sec:results} presents a summary of our results on the
tree path labeling problem.

@@

As described earlier, a matrix with COP is one whose rows (columns)
can be rearranged so that the 1s in every column (row) are in
consecutive rows (columns).

@@

\cite{bl76} describes a linear time construction of a data structure
called PQ trees. This paper draws on the close relationship of COP
matrices to interval graphs. A PQ tree of a matrix is one that stores
all row (column) permutations of the matrix that give the COP orders
(there could be multiple) of the matrix. This is constructed using an
elaborate linear time procedure. Since this result, several
improvements have been made to the procedural complexity of the
algorithm itself and also derivative data structures that are easier
to construct. PQR trees is one such data structure \cite{mm96},
\cite{mpt98}\tnote[TH]{also, mpt98 $==$ elaborate version of mm96}.
\cite{tm05} describes an improved version of PQR
trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
simpler algorithm for COT. Hsu also invented PC trees
\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
which is claimed to be much easier to implement. \cite{nsnrs09}
describes a characterization of consecutive ones property solely based
on the cardinality properties of the set representations of the
columns (rows); every column (row) is equivalent to a set that has the
row (column) indices of the rows (columns) that have one entries in
this column (row). This is interesting and relevant, especially to
this thesis because it simplifies COT to a great
degree. \tnote{get this
  figured out! - because it reduces the search space for the solution
  from set of all permutations of rows to set of all}

\cite{mcc04} describes a different approach to COT. While all previous
COT algorithms gave the COP order if the matrix has the property and
if not it exited stating negative, this algorithm gives an evidence by
way of a certificate of
matrix being devoid of COP thereby enabling a user to check the
algorithm's result. This is significant from an implementation
perspective because automated program verification is hard and manual
verification is more viable. Hence having a certificate
reinforces its credibility. Note that when the matrix has COP the COP
order is the certificate.  It may also be noted that the internal
machinery of this algorithm \annote {is related to the weighted
  betweenness problem addressed } {in what way??} in \cite{co98}.
\tnote{expand on the COP order graph creation and it having to be
  bipartite for M to have COP. and thus an odd cycle being an evidence
  of no COP.}
\tnote[TH]{ where should this go?: cite|jlm97 (application of PQ trees
  in graphics).\\ add helly's theorem citation
  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}

The literature also has results on parallel COT algorithms --
\cite{as95} \cite{bs03} \cite{ly91}\tnote{add english to fit context}

\subsection{Optimization problems in COP}
So far we have been concerned about
matrices that have the consecutive ones property. However in real life
applications, it is rare that data sets represented by binary matrices
have COP. Nevertheless, the COP is not arbirary and there are several
interesting problems when a matrix does not have COP but is ``close''
to having COP or can be altered to have COP.

-- sect 4.1 in \cite{d08phd} has many results surveyed. hardness results,
approx. results. results are usually for a class of matrices $(a,b)$
where number columns and rows are restriced to $a$ and $b$
\tnote{check if it is the other way around}.\\
-- problem of flipping at most $k$ entries of $M$ to make it attain
COP. this is NP complete \cite{b75-phd}

The first step in this direction is to analyze matrices without
COP. \cite{at72} showed that a matrix that does not have COP have
certain forbidden substructures that prevent it from having
COP. Tucker classified these substructures into five classes of
submatrices. This result is presented in the context of AT-free
bipartite graphs \tnote[TH]{ check this up. give details.} which is
equivalent to matrices with COP. A binary matrix is converted into a
bipartite graph as follows. The set of rows and the set of columns
form the two partitions of the graph. Each row node is adjacent to
those nodes that represent the columns that have ones in the
corresponding row. \cite{at72} proves that this bipartite graph has no
asteroidal triple if and only if the matrix has COP and goes on to
identify the forbidden substructures for these bipartite graphs. The
matrices corresponding to these substructures are the forbidden
submatrices. Once a matrix has been detected to not have COP (using
any of the COT algorithms mentioned earlier), it is naturally of
interest to find out the smallest (in terms of number of rows, columns
and number of entries that are ones) forbidden
substructure. \cite{d08phd} discusses an exact algorithm with time
complexity \annote{parameterized by the number of ones in the
  matrix}{check this.}.

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete\tnote{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hg02} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem\tnote[TH]{elaborate - what are
  the results?}.

Another problem is to find the minumum number of entries in the matrix
that can be flipped to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to ones so that the
resulting matrix has COP. As mentioned earlier, this problem had been
known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
using a reduction to the longest path problem, \tnote{or is it a
  survey of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$ rows is NP complete. \tnote{how is this
  different from booth's 75 result??}  \tnote[TH]{where should this
  go? cite|tz04 (approx submatrix with COP sparse matrices)}

\cite{jkckv04} discusses the use of matrices with almost-COP in the
storage of very large databases.  The problem is that of $k$-run
reordering of binary matrices which they prove to be NP hard using a
reduction from the hamiltonian path problem.  A connection of COP
problem to the travelling salesman problem is also
introduced. \tnote[TH]{what does this mean? -- COP can be used as a
  tool to reorder $0.5T \le runs(M) \le T$}

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{k77}. \tnote[TH]{what is the reduction?}

\tnote[TH]{other problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

{\bf Motivation: } A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

Our results also have close connection to recognition of path graphs
and connection to path graph isomorphism.  Graphs which can be
represented as the intersection graph of paths in a tree are called
{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
be viewed as paths in a tree, if and only if the intersection graph of
$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
graphs since chordal graphs are combinatorially characterized as the
intersection graphs of subtrees of a tree.  Path graphs are well
studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
a graph is a path graph can be done in polynomial time by the results
of \cite{gav78,aas93}.  However, this is only a necessary condition
for our question.  \remove{Indeed, our work provides a sufficient
  condition for a related problem: that is whether a path assignment
  to be feasible?}  On the other hand, path graph isomorphism is known
be isomorphism-complete; see for example \cite{kklv10}. Therefore,
\annote{it is unlikely that we can solve the problem of finding
  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
\annote{It is definitely interesting to classify the kinds of trees
  and hypergraphs for which feasible path labelings can be
}{ELABORATE.} found efficiently.  \annote{These results would form
  a natural generalization of COP testing and interval graph
  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.

@@

\section{Summary of results}
\label{sec:results}

@@

%\item 
In the Section 4, \annote{we initiate an exploration of}{WTF?}
finding feasible path labeling of set systems in a special kind of
tree which we call the $k$-subdivided star.  \annote{This question
  is an attempt to generalize}{REWORD.} the problem of testing if a
matrix has the consecutive ones property.  However, we restrict the
hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
elements.  \annote{In spite of this restricted case we consider, we
  believe that our results are of significant interest in
  understanding the nature of Graph Isomorphism}{REWORD.} which is
polynomial time solvable in interval graphs and is hard on path
graphs.
%\end{enumerate}

@@

%\subsection{RESULTS}
% \section{Preliminaries} 
% \label{sec:prelims}

@@

\subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
\label{sec:feasible}

In this section we give an algorithmic characterization of a
feasibility of tree path labeling.
% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% and path system $\cP$ on the given tree $T$. We
% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
%   (ICPPL)} if it has the following properties.
Consider a path labeling $(\cF, \cl)$ on the
given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
  Preserving Path Labeling (ICPPL)} if it has the following
properties.

% \begin{enumerate}[{\icpplpr\ }i. ]
% \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
%   for all distinct $S_1, S_2 \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
%   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
% \end{enumerate}

\begin{table}[h]
  \centering
  \begin{tabular}[l]{rrcll}
    Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
    Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
    \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
    Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
    |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
    S_2, S_3 \in \cF$ 
  \end{tabular}
  \caption{ICPPL properties}
  \label{tab:icpplpr}
\end{table}

\noindent
The following lemma is useful in subsequent arguments. 
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
